Continuous Deployment and Schema Evolution \\in SQL Databases

### 2015-01-29 paper.tex
Fixed minor typos and rephrased some sentences.

Continuous Deployment is an important enabler rapid delivery of business value and early end user feedback. While frequent code deployment is well understood, the impact of frequent change on persistent data is less understood and supported. SQL schema evolutions in particular can make it expensive to deploy a new version, and may even lead to downtime if schema changes can only be applied by blocking operations.

In this paper we study the problem of continuous deployment in the presence of database schema evolution in more detail. We identify a number of shortcomings to existing solutions and tools, mostly related to avoidable down time and support for foreign keys. We propose a novel approach to address these problems, and provide an open source implementation. Initial evaluation suggests the approach is effective and sufficiently efficient.

@@

When service windows are short or even non-existent, these two problems make releasing new versions of applications frequently problematic for release engineers. To address this problem we aim to find or implement a tool which  allows its users to not only alter their database schema in a non-blocking fashion, but also do this without affecting older and still running applications.

@@

\subsubsection{SoundCloud's Large Hadron Migrator~\cite{SoundcloudLHM} \& Facebook's Online Schema Change~\cite{FacebookOSC}}
These are both tools which closely follow Approach~1 descriged earlier (Section~\ref{sec:cd}). They create a copy of a table (including its contents) and record modifications to records in the original table using triggers into a special ``deltas'' table. Once they have copied the original table and its contents, they apply the specified schema operations to the ghost table before proceeding to replay the recorded modifications from the ``deltas'' table onto the ghost table. When the ``deltas'' table is almost empty, these tools acquire an exclusive write lock on the original table --- like \textit{TableMigrator} does --- and apply the last of the recorded modifications before atomically switch the names of both tables.

@@

\begin{itemize}
  \item None of the tools used in practice support foreign keys
  \item The tools primarily focus on MySQL databases.
  \item These tools are limited in the sense that they typically only support making small changes to the existing database schema. This forces software and release engineers to often release intermediate versions of the application, as the schema is altered in several steps.
\end{itemize}

@@

It is worth noting that we have not invested any time in improving the performance of QuantumDB or reducing the performance impact it has on the database and its clients. During this experiment we noticed that the hard disk was periodically unable to write data to disk quickly enough. As a result queries were taking significantly longer than normal to complete. We are confident that we can greatly reduce this by tweaking the PostgreSQL config, or switching to a faster hard disk.

>>>>

Continuous Deployment is an important enabler of rapid delivery of business value and early end user feedback. While frequent code deployment is well understood, the impact of frequent change on persistent data is less understood and supported. SQL schema evolutions in particular can make it expensive to deploy a new version, and may even lead to downtime if schema changes can only be applied by blocking operations.

In this paper we study the problem of continuous deployment in the presence of database schema evolution in more detail. We identify a number of shortcomings to existing solutions and tools, mostly related to avoidable downtime and support for foreign keys. We propose a novel approach to address these problems, and provide an open source implementation. Initial evaluation suggests the approach is effective and sufficiently efficient.

@@

When service windows are short or even non-existent, these two problems make releasing new versions of applications frequently problematic for release engineers. To address this problem we aim to find or implement a tool which allows its users to not only alter their database schema in a non-blocking fashion, but also do this without affecting older applications still operating on the older version of the schema.

@@

\subsubsection{SoundCloud's Large Hadron Migrator~\cite{SoundcloudLHM} \& Facebook's Online Schema Change~\cite{FacebookOSC}}
These are both tools which closely follow Approach~1 described earlier (Section~\ref{sec:cd}). They create a copy of a table (including its contents) and record modifications to records in the original table using triggers into a special ``deltas'' table. Once they have copied the original table and its contents, they apply the specified schema operations to the ghost table before proceeding to replay the recorded modifications from the ``deltas'' table onto the ghost table. When the ``deltas'' table is almost empty, these tools acquire an exclusive write lock on the original table --- like \textit{TableMigrator} does --- and apply the last of the recorded modifications before atomically switch the names of both tables.

@@

\begin{itemize}
  \item None of the tools used in practice support foreign keys.
  \item The tools primarily focus on MySQL databases.
  \item These tools are limited in the sense that they typically only support making small changes to the existing database schema. This forces software and release engineers to often release intermediate versions of the application, as the schema is altered in several steps.
\end{itemize}

@@

It is worth noting that we have not invested any time in improving the performance of QuantumDB or reducing the performance impact it has on the database and its clients. During this experiment we noticed that the hard disk was periodically unable to write data to disk quickly enough. As a result queries were taking significantly longer than normal to complete. We are confident that we can greatly reduce this by tweaking the PostgreSQL config, switching to a faster hard disk, or lowering the rate of migration during the copy phase.
